---
title: "Palmer Penguins Explorer"
subtitle: "Exploring penguin measurements across species and islands in Antarctica."
---

::: {.callout-note}
## Learning Objectives
- Visualize relationships between penguin body measurements
- Compare species distributions across islands using plots and tables
- Explore grouped summaries with interactive tabsets
:::

```{r}
#| label: setup
#| message: false
#| warning: false

library(palmerpenguins)
library(ggplot2)
library(dplyr)
library(gt)
```

## Overview

The [Palmer Penguins](https://allisonhorst.github.io/palmerpenguins/) dataset contains measurements for `r nrow(penguins)` penguins from three species observed on three islands in the Palmer Archipelago, Antarctica.

::: {.gated}
## Meet the Species

Click a card to learn more about each species.

::: {.grid}

::: {.g-col-12 .g-col-md-4}
<div class="flip-card flip-card-adelie" onclick="this.classList.toggle('flipped')">
<div class="flip-card-inner">
<div class="flip-card-front">
<h3>Adelie</h3>
<p>Click to flip</p>
</div>
<div class="flip-card-back">
<p><strong>Islands:</strong> All three</p>
<p><strong>Avg mass:</strong> 3,701 g</p>
<p><strong>Avg flipper:</strong> 190 mm</p>
<p><strong>Diet:</strong> Krill</p>
</div>
</div>
</div>
:::

::: {.g-col-12 .g-col-md-4}
<div class="flip-card flip-card-chinstrap" onclick="this.classList.toggle('flipped')">
<div class="flip-card-inner">
<div class="flip-card-front">
<h3>Chinstrap</h3>
<p>Click to flip</p>
</div>
<div class="flip-card-back">
<p><strong>Island:</strong> Dream</p>
<p><strong>Avg mass:</strong> 3,733 g</p>
<p><strong>Avg flipper:</strong> 196 mm</p>
<p><strong>Diet:</strong> Krill</p>
</div>
</div>
</div>
:::

::: {.g-col-12 .g-col-md-4}
<div class="flip-card flip-card-gentoo" onclick="this.classList.toggle('flipped')">
<div class="flip-card-inner">
<div class="flip-card-front">
<h3>Gentoo</h3>
<p>Click to flip</p>
</div>
<div class="flip-card-back">
<p><strong>Island:</strong> Biscoe</p>
<p><strong>Avg mass:</strong> 5,076 g</p>
<p><strong>Avg flipper:</strong> 217 mm</p>
<p><strong>Diet:</strong> Fish & krill</p>
</div>
</div>
</div>
:::

:::
:::

::: {.gated}
## Scatter Plot

```{r}
#| label: fig-scatter
#| fig-cap: "Flipper length vs. body mass by species"
#| warning: false

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species, shape = species)) +
  geom_point(size = 2.5, alpha = 0.7) +
  scale_color_manual(values = c("darkorange", "purple", "cyan4")) +
  labs(
    x = "Flipper Length (mm)",
    y = "Body Mass (g)",
    color = "Species",
    shape = "Species"
  ) +
  theme_minimal()
```
:::

::: {.gated}
## Summary Table

```{r}
#| label: tbl-summary

penguins |>
  filter(!is.na(body_mass_g)) |>
  group_by(species) |>
  summarise(
    Count = n(),
    `Mean Body Mass (g)` = round(mean(body_mass_g), 0),
    `Mean Flipper Length (mm)` = round(mean(flipper_length_mm, na.rm = TRUE), 1),
    `Mean Bill Length (mm)` = round(mean(bill_length_mm, na.rm = TRUE), 1),
    `Mean Bill Depth (mm)` = round(mean(bill_depth_mm, na.rm = TRUE), 1)
  ) |>
  gt() |>
  tab_header(
    title = "Penguin Species Summary",
    subtitle = "Averages across all islands"
  ) |>
  cols_label(species = "Species") |>
  tab_source_note("Source: palmerpenguins R package")
```
:::

::: {.gated}
## Did You Know?

Swipe or use the arrows to browse penguin fun facts.

<div class="slider-wrapper">
<button class="slider-btn slider-btn-prev" onclick="this.parentElement.querySelector('.slider-track').scrollBy({left: -296, behavior: 'smooth'})">&#8249;</button>
<div class="slider-track">

<div class="slider-card">
<h4>Deepest Divers</h4>
<p>Gentoo penguins can dive to depths of over 200 meters and hold their breath for up to 7 minutes while hunting fish.</p>
</div>

<div class="slider-card">
<h4>Speed Swimmers</h4>
<p>Gentoos are the fastest underwater-swimming penguins, reaching speeds of up to 36 km/h (22 mph).</p>
</div>

<div class="slider-card">
<h4>Pebble Proposals</h4>
<p>Adelie penguins propose to their mates by offering a carefully chosen pebble. Fights can break out over the best stones.</p>
</div>

<div class="slider-card">
<h4>Named for a Stripe</h4>
<p>Chinstrap penguins get their name from the narrow black band running under their chin, like a helmet strap.</p>
</div>

<div class="slider-card">
<h4>Extreme Commuters</h4>
<p>Adelie penguins migrate up to 17,600 km (10,900 mi) each year, following the sun between breeding and feeding grounds.</p>
</div>

<div class="slider-card">
<h4>Largest of the Three</h4>
<p>Gentoo penguins are the largest species in this dataset, averaging over 5 kg — about 37% heavier than Adelie or Chinstrap.</p>
</div>

</div>
<button class="slider-btn slider-btn-next" onclick="this.parentElement.querySelector('.slider-track').scrollBy({left: 296, behavior: 'smooth'})">&#8250;</button>
</div>
:::

::: {.gated}
## Details by Island

::: {.panel-tabset}

### Biscoe

```{r}
#| warning: false
penguins |>
  filter(island == "Biscoe") |>
  ggplot(aes(x = species, y = body_mass_g, fill = species)) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  scale_fill_manual(values = c("darkorange", "cyan4")) +
  labs(x = NULL, y = "Body Mass (g)") +
  theme_minimal()
```

### Dream

```{r}
#| warning: false
penguins |>
  filter(island == "Dream") |>
  ggplot(aes(x = species, y = body_mass_g, fill = species)) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  scale_fill_manual(values = c("darkorange", "purple")) +
  labs(x = NULL, y = "Body Mass (g)") +
  theme_minimal()
```

### Torgersen

```{r}
#| warning: false
penguins |>
  filter(island == "Torgersen") |>
  ggplot(aes(x = species, y = body_mass_g, fill = species)) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  scale_fill_manual(values = c("darkorange")) +
  labs(x = NULL, y = "Body Mass (g)") +
  theme_minimal()
```

:::
:::

::: {.gated}
## About the Data

::: {.callout-tip collapse="true"}
## Data Collection Details

The Palmer Penguins dataset was collected by [Dr. Kristen Gorman](https://www.uaf.edu/cfos/people/faculty/detail/kristen-gorman.php) and the [Palmer Station, Antarctica LTER](https://pallter.marine.rutgers.edu/), a member of the Long Term Ecological Research Network. The data were collected from 2007--2009 and include the following variables:

- **species**: Adelie, Chinstrap, or Gentoo
- **island**: Biscoe, Dream, or Torgersen
- **bill_length_mm**: Length of the bill (mm)
- **bill_depth_mm**: Depth of the bill (mm)
- **flipper_length_mm**: Flipper length (mm)
- **body_mass_g**: Body mass (g)
- **sex**: Male or Female
- **year**: Study year (2007, 2008, or 2009)

This dataset is commonly used as an alternative to the `iris` dataset for data exploration and visualization examples.
:::

::: {.callout-tip collapse="true"}
## Missing Values

```{r}
penguins |>
  summarise(across(everything(), ~ sum(is.na(.)))) |>
  gt() |>
  tab_header(title = "Missing Values per Column")
```
:::
:::
