---
title: "Palmer Penguins Explorer"
---

::: {.lesson-hero}
# Palmer Penguins Explorer
::: {.lead}
Exploring penguin measurements across species and islands in Antarctica.
:::
<span class="badge-feature badge-static">ggplot2</span>
<span class="badge-feature badge-static">gt tables</span>
<span class="badge-feature badge-static">Tabsets</span>
:::

::: {.learning-objectives}
#### Learning Objectives
- Visualize relationships between penguin body measurements
- Compare species distributions across islands using plots and tables
- Explore grouped summaries with interactive tabsets
:::

```{r}
#| label: setup
#| message: false
#| warning: false

library(palmerpenguins)
library(ggplot2)
library(dplyr)
library(gt)
```

## Overview

The [Palmer Penguins](https://allisonhorst.github.io/palmerpenguins/) dataset contains measurements for `r nrow(penguins)` penguins from three species observed on three islands in the Palmer Archipelago, Antarctica.

<hr class="section-divider">

## Scatter Plot

<div class="card">
<div class="card-header">
Flipper Length vs. Body Mass
</div>
<div class="card-body">

```{r}
#| label: fig-scatter
#| fig-cap: "Flipper length vs. body mass by species"
#| warning: false

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species, shape = species)) +
  geom_point(size = 2.5, alpha = 0.7) +
  scale_color_manual(values = c("darkorange", "purple", "cyan4")) +
  labs(
    x = "Flipper Length (mm)",
    y = "Body Mass (g)",
    color = "Species",
    shape = "Species"
  ) +
  theme_minimal()
```

</div>
</div>

<hr class="section-divider">

## Summary Table

<div class="card">
<div class="card-header">
Species Averages
</div>
<div class="card-body">

```{r}
#| label: tbl-summary

penguins |>
  filter(!is.na(body_mass_g)) |>
  group_by(species) |>
  summarise(
    Count = n(),
    `Mean Body Mass (g)` = round(mean(body_mass_g), 0),
    `Mean Flipper Length (mm)` = round(mean(flipper_length_mm, na.rm = TRUE), 1),
    `Mean Bill Length (mm)` = round(mean(bill_length_mm, na.rm = TRUE), 1),
    `Mean Bill Depth (mm)` = round(mean(bill_depth_mm, na.rm = TRUE), 1)
  ) |>
  gt() |>
  tab_header(
    title = "Penguin Species Summary",
    subtitle = "Averages across all islands"
  ) |>
  cols_label(species = "Species") |>
  tab_source_note("Source: palmerpenguins R package")
```

</div>
</div>

<hr class="section-divider">

## Details by Island

::: {.panel-tabset}

### Biscoe

```{r}
#| warning: false
penguins |>
  filter(island == "Biscoe") |>
  ggplot(aes(x = species, y = body_mass_g, fill = species)) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  scale_fill_manual(values = c("darkorange", "cyan4")) +
  labs(x = NULL, y = "Body Mass (g)") +
  theme_minimal()
```

### Dream

```{r}
#| warning: false
penguins |>
  filter(island == "Dream") |>
  ggplot(aes(x = species, y = body_mass_g, fill = species)) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  scale_fill_manual(values = c("darkorange", "purple")) +
  labs(x = NULL, y = "Body Mass (g)") +
  theme_minimal()
```

### Torgersen

```{r}
#| warning: false
penguins |>
  filter(island == "Torgersen") |>
  ggplot(aes(x = species, y = body_mass_g, fill = species)) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  scale_fill_manual(values = c("darkorange")) +
  labs(x = NULL, y = "Body Mass (g)") +
  theme_minimal()
```

:::

<hr class="section-divider">

## About the Data

<div class="accordion" id="aboutAccordion">
<div class="accordion-item">
<h2 class="accordion-header">
<button class="accordion-button collapsed" type="button" data-bs-toggle="collapse" data-bs-target="#collapseCollection">
Data Collection Details
</button>
</h2>
<div id="collapseCollection" class="accordion-collapse collapse" data-bs-parent="#aboutAccordion">
<div class="accordion-body">

The Palmer Penguins dataset was collected by [Dr. Kristen Gorman](https://www.uaf.edu/cfos/people/faculty/detail/kristen-gorman.php) and the [Palmer Station, Antarctica LTER](https://pallter.marine.rutgers.edu/), a member of the Long Term Ecological Research Network. The data were collected from 2007--2009 and include the following variables:

- **species**: Adelie, Chinstrap, or Gentoo
- **island**: Biscoe, Dream, or Torgersen
- **bill_length_mm**: Length of the bill (mm)
- **bill_depth_mm**: Depth of the bill (mm)
- **flipper_length_mm**: Flipper length (mm)
- **body_mass_g**: Body mass (g)
- **sex**: Male or Female
- **year**: Study year (2007, 2008, or 2009)

This dataset is commonly used as an alternative to the `iris` dataset for data exploration and visualization examples.

</div>
</div>
</div>
<div class="accordion-item">
<h2 class="accordion-header">
<button class="accordion-button collapsed" type="button" data-bs-toggle="collapse" data-bs-target="#collapseMissing">
Missing Values
</button>
</h2>
<div id="collapseMissing" class="accordion-collapse collapse" data-bs-parent="#aboutAccordion">
<div class="accordion-body">

```{r}
penguins |>
  summarise(across(everything(), ~ sum(is.na(.)))) |>
  gt() |>
  tab_header(title = "Missing Values per Column")
```

</div>
</div>
</div>
</div>
